TO FIND

### 2013-05-27 ms.tex
First complete reading

% TODO:
% TP, PDP & DG: Complete the discussion
% PDP: Additional paragraph on the metrics
% DG: Figure legends
% DG: Abstract
% DG: Compile references
% TP & PDP: fix bugs I generated!!!

@@

Ecology needs to move toward a more predictive approach. Integrating
elements of theoretical ecology \parencite{Thuiller2013} with the metacommunity
perspective appears naturally as a good candidate to fill this challenge. The metacommunity
concept builds on feedbacks between local scale processes, such as competitive
interactions and local adaptation, and regional scale processes such as
dispersal, gene flow and speciation \textcite{Leibold2004a}. This perspective is
particularly relevant to limnology, where exchanges of organisms and nutrients
affect community and ecosystem properties from the local (e.g. vertical mixing
\parencite{Ryabov2011}) to the regional (e.g. connection of lakes
\parencite{Leibold2004b} scales. It emphasizes the importance of dispersal
relative to pairwise interactions in the organization of ecological communities.

At first sight, the development of neutral theory appears as a step back. Based on Kimura's neutral theory of molecular evolution, the neutral theory of biodiversity makes the provocative
assumption that species are ecologically equivalent. Thus, variation in
the environment has no impact on demography \parencite{Bell2000,Hubbell2001}. Neutral
ecological communities are driven only by demographic stochasticity and dispersal. Neutral theory sparked an historical debate still lasting after more
than a decade \parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It
was stimulated by the surprising ability of neutral models to fit some well
studied empirical observations such as species abundance distributions and
distance-decay relationships. A remarkable strength of the theory is to
provide a \emph{"formal general theory of abundance and diversity that will
account, in a simple and economical fashion, for the many patterns that
ecologists have documented"} \parencite{Bell2001}. Even if new studies rejecting
neutral theory are consistently published (e.g.  \textcite{Ricklefs2012}), a consensus is forming that neutral theory is a well-developed null
hypothesis for niche theory and could even be used as an adequate approximation
of ecological dynamics in some situations.  \textcite{Bell2001} nicely
envisioned two perspectives to neutral theory that are still standing today.
Under the weak perspective, neutral theory provides a set of realistic
predictions of community organization despite false assumptions. Even if being
fundamentally wrong, neutral theory is still useful when used as a null
hypothesis \parencite{Gotelli2006}. It is considered as an improvement over
traditional null hypotheses based on randomization \parencite{Gotelli2000}
because it readily integrates dispersal. The strong version on the other hand
posits that neutral theory is a satisfying approximation to community dynamics
and an appropriate theory to explain the distribution of biodiversity. It
implies that the right mechanisms have been identified and that the consistently
observed differences among species do not impact community organization.

Neutral theory has also been proposed as a useful tool to understand and predict
some aspects of community dynamics. It links to an old philosophical debate
between realism and instrumentalism \parencite{Wennekes2011}. Because every
ecological model is a simplification of reality, scientists have to
subjectively decide the level of details they put in, leaving out some elements they consider unimportant. The realism perspective requires that all assumptions of
theory must be true, while the utility of the theory is more important to
instrumentalism. The utilitarian value of a theory could either be for
understanding or prediction (another old philosophical debate, see
\textcite{Schmueli2010}). Obviously neutral theory could only be instrumental.
The question then is if such a 'general, large-scale, but vague' theory
\parencite{Wennekes2011} is a satisfying approximation.

The instrumentalist view of neutral theory raises the question of why it should
be a satisfying approximation despite knowing the pieces are wrong? Perhaps stochasticity
of various origins blur the deterministic differences among species and
promote ecological drift \parencite{Gravel2011}. A second answer is that dispersal and historical contingencies might
have a much more profound impact on species distribution \parencite{Bahn2007,
Boulangeat2012} and ecological dynamics. The debate over the equivalence
assumption and demographic stochasticity has perhaps overlook the recognition of
how much dispersal influence community assembly.

@@

A network is a discrete mathematical object made of two sets: a set of
vertices (or nodes) $V$ and a set of edges $E$ connecting the vertices
\cite{new10}. The term ``graph'' is often preferred in computer science and
mathematics \cite{gro06}, with graph algorithms being an important and active
area of research \cite{sed01}. A network is a combinatorial object: it is
used to study how discrete entities are connected and how they combine
together to create complex structures. They are used to study molecules, food
webs, social networks, or even the relationship between variables in
statistics \cite{wri21,new10}. We are especially interested in spatial
networks, a special kind of network mixing the combinatorial properties of
networks with a topological space \cite{kob94}. Thus, the vertices in a
spatial graph are embedded in some other space, most often the two or three-
dimensional Euclidean space. This object brings a rich representation to
spatial ecology and is particularly suited for systems of lakes and rivers, which can
easily be represented by vertices and edges. There are two notions of
distance in spatial networks. Euclidean distance represents the geographical
distance between the vertices $(i, j)$, i.e.: $\sqrt{(x_i - x_j)^2 + (y_i - y_
j)^2}$. Geodesic distance is the distance in the graph space, i.e.: the
length of the shortest path \cite{dij59}. For example, two lakes could be
very close on a map (short Euclidean distance) but the geodesic distance could be great if they are not directly
linked by a river.

The popularity of network theory stem for their ability to model complex
structures while allowing us to extract useful metrics (Table 1). At a very high level, 
a network can be described by a number of vertices (the order) and a number
of edges (the size). Looking more closely, the relationship between vertices
is influenced by paths, which are ordered series of nodes.
Centrality is a central (no pun intended) concept in network theory, where it
can be seen to as a measure of ``importance''. The simplest measure
of the centrality of a node is its degree, which is the number of nodes directly connected to it. Of course, this is a very rough description of centrality. For
example, two lakes can have the same degree, with one being connected to
a small isolated cluster, while the other one is part of one of the biggest 
network of lake. In this case, measures of centrality like eigen-centrality will
weight the importance of the connection so a node connected to well-connected
nodes will have higher centrality than a node connected to isolated nodes.

For simulations, spatial networks can easily be generated with the random
geometric graph algorithm \cite{sed01}. In this algorithm, all vertices are
assigned to a position in some two-dimensional space, most often the unit
square. Then, all pairs of vertices within some threshold Euclidean distance $r$
are connected with an edge. The resulting networks have the desirable
property of locality: if a vertex $A$ is connected to two vertices $B$ and $C$.
then $B$ and $C$ are more likely to be connected than two random vertices.
Random geometric networks have been extensively studied \cite{app97a,app97
b,app02a,app02b,pen03} and we provide a R function to generate them. The 
position of vertices is typically random, but we could also imagine alterations
where they are either more aggregated or segregated than expected by chance alone.

We also provide the code for a second structure that we call a random geometric
tree.  It builds a tree from the shortest path tree \cite{dij59} of a vertex in
a random geometric network. This random geometric tree does not exactly
represent dendritic landscapes but is a convenient model to simulate lake
connected by rivers to a series of smaller lakes.

@@

% Table 1: main descriptors of spatial networks. Column 1: Name of the metric; Column 2: Definition
% We might want to distinguish network level versus node-level metrics.
Concepts
Path A sequence of edges forming a sequence of vertices.
Connection Two vertices are connected iff there is a path between them.
Geodesic distance Length of the shortest path between two vertices.
Network-level metrics
Order Total number of vertices.
Size Total number of edges.
Connectivity A measure of robustness: the minimum number of elements to remove
to isolate the vertices.
Components The number of connected subsets.
Vertex-level metrics
Degree The number of edges of a vertex.
Closeness centrality Average geodesic distance between a vertex and all other vertex.
Eigenvector centrality A measure of centrality based on the concept that connection to
highly connected vertices are more important.
Betweenness centrality The number of shortest paths from all vertices to all others that
pass through that vertex.

@@

In this section we describe three toy models representing different
perspectives of metacommunity ecology: patch dynamics, neutral dynamics and
species sorting. While the neutral model is interesting in itself, it is by
comparison with a model without any interactions (patch dynamics) and with niche
differentiation (species sorting) that we will be able to fully understand the
interaction between these processes and landscape structure. Despite neutral,
competitive interactions in neutral models are very strong because of the
zero-sum assumption (the community is always at carrying capacity). We will
first review the fundamental assumptions of each model with their description
(Table 2 summarizes the parameters and variables that are used), and then
briefly discuss their main predictions. Simulation results are presented in the
next section, with the corresponding R code provided in the Supplementary
Material.

@@

The simplest metacommunity model is a $S$ species extension of traditional
metapopulation models \parencite(Hanski1999). The standard Levins metapopulation
model \parencite(Levins1969) describes the stochastic colonizations and
extinctions of population over a homogenous landscape. The basic unit is the
population. The Levins model tracts the dynamics of occupancy (the fraction of
the landscape that is occupied) with an ordinary differential equation and
therefore assumes an infinite landscape. The simulation model we run is more
realistic as it simulates a finite number $N$ of discrete patches. The rules
described in the previous section were used to generate connectivity matrices
along four scenarios (Fig. 1): global dispersal (connected graph), a lattice, a
random geometric graph and a random tree graph. A patch $x$ shares $d_x$ links
with neigbhouring patches (its degree). At each time step (the simulation model
is discrete in time), the probability that a colonist coming from an occupied
patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$ is the probability a
colonization event takes place if all connected patches are occupied. The
expected probability that a colonist arrives to patch $x$ from patch $y$ is then
$C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability that patch $y$ is
occupied by species $i$. The probability that an extinction occurs in a given
patch is $e$. The Levins model is for a single species, but a basic
metacommunity patch dynamics model could be run by aggregating $S$ independent
metapopulation models \parencite(Hanski1997). There are no interactions in this
simple model, which means there is no limit to local species richness and no
carrying capacity. Competitive, mutualistic and predator-prey interactions have
been added to this framework (e.g.
\textcite{Tilman1994,Klausmeier1998,Holt1996}) but we will keep this model
minimal for the sake of comparison with the neutral model.

Predictions of the patch dynamics metacommunity model are quite straightforward.
First, a fundamental result of metapopulation ecology is that persistence will
occur if colonization probability is larger than extinction probability ($c>e$).
Given that all species are the same, then we should expect the regional
diversity to be $S$ if this condition is satisfied and $0$ if not. The situation
is however more complex in spatially explicit landscapes with complex
connectivity matrices \parencite{Hanski1998}. Spatially explicit dispersal
usually reduces the occupancy and thereby the likelihood of persistence. The
second prediction is that, given spatial variation in connectivity, there will
be spatial variation in occurrence probability. Given the above formulation of a
colonization event to occur, the probability that an empty location is colonized
is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation basically tells us that the
colonization probability will increase asymptotically with the degree of a patch
(because of the product). It is easy to show from metapopulation theory that the
occurrence probability in a patch is then $p_ix=I_x(I_ix+e)^{-1}$. The feedback
between local and regional dynamics arises because all $p_ix$ from the landscape
are dependent from each other. Simulations are usually conducted to solve the
model for a large landscape, but numerical solutions are theoretically possible.
The aggregation across the $S$ species of the regional species pool is obtained
by taking the summation of occurrence probabilities over all species, $s_x =
\sum{p_i}$. Because in this model all species are equal, we expect the local
species richness to be a linear function of the patch degree (number of edges). Multi-species
analysis of metapopulation models also reveals interesting predictions on other
aspects of community organization at various spatial scales such as the species-area
relationship \parencite{Hanski1997}, and proved to be useful in conservation
ecology with predictions of extinctions following habitat destruction
\parencite{Nee1994,Rybicki2013}.

@@

The formulation of the recruitment probability is the central piece of all
neutral models, making possible the coupling with the metacommunity and
neighbouring patches. We adopt a simple formulation in our model based on
\parencite{Gravel2006}. The approach is conceptually similar to placing a seed
trap in a canopy gap and picking a seed at random among the ones falling
in to determine the identity of the recruited species. The composition of the
seed pool in that trap will be a mixture of local dispersal and immigrants from
the metacommunity. For simplicity, we consider three spatial scales of dispersal
but it would be easy to generalize the approach to a continuous seed dispersal
kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
recruit is a migrant from neighbouring patches, $M$ is the probability it comes
from a larger (and fixed) metacommunity, and consequently, by substraction,
$1-m-M$ is the probability it comes from local dispersal. The fraction
$N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative
abundance of species $i$ in the seed pool coming from neigbouring patches $x$.
The relative abundance in the neighborhood is weighted by the degree of the
connected nodes because some nodes will spread their seeds across a higher
number of nodes and thus contribute less to the seed pool. We consider simply
$P_{ix} =\frac{\sum d_y^{-1}P_{iy}}{\sum dy^{-1}}$.  We assume for simplicity
(and without loss of generality, \parencite{Bell2000}) that the relative
abundance in the metacommunity is uniform, i.-e. equal to $S^{-1}$. This
immigration prevents the collapse of the metacommunity because otherwise all
species except one will face extinction by ecological drift (speciation prevents
this phenomenon to occur in \textcite{Hubbell2001}). The local recruitment
probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
(1-m-M)N_{ix}J_x^{-1}$.

The model is neutral because it assumes that the probabilities of a local
recruitment, an immigration and a mortality event are all equal across species. Demographic
stochasticity is the source of variations in abundance, but larger disturbances
could be simulated as well, as long as they hit all species with the same
probability, independently of their density. The fundamental feature of neutral
dynamics is therefore the ecological drift, defined as population changes
emerging from neutraly stable population dynamics. It can be measured as the
variance between replicated time series of community dynamics
\parencite{Gravel2011}. \textcite{Hubbell2001} provides a very comprehensive
analysis of the model, with specific attention to the effect of the different
parameters on the drift (and consequently variance in abundance) and time to
extinction. Despite its simplicity, the neutral model is surprisingly rich in
the predictions it makes. \textcite{Bell2001} and \textcite{Hubbell2001}
analyzed the performance of neutral models to predict species abundance
distributions, range-abundance relationship, spatial variation in abundance,
species-area relationship, community turnover (beta-diversity) and
co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parenciteCanard2012}. Other than the ecological equivalence assumption, one of the most
criticized aspect of neutral models is the realism of the speciation process and
the required speciation rates to sustain species richness
\parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more credible
speciation models \parencite{Rosindell2009,Desjardins2012a} revealed the
difficulty to maintain diversity in neutral models over macro-evolutionary time
scales. These models however also generated interesting novel predictions on
endemic species richness \parencite{Rosindell2011,Desjardins2012b}.

@@

The species-sorting and the mass effect perspectives build on the notion of
species-specific responses to a spatially varying environment
\parencite{Leibold2004a}. There are various ways to simulate such dynamics and
we picked the lottery model, in line with tradition \parencite{Mouquet2002} and
for its proximity to the neutral model described above. Competition for space
occurs during recruitment after the death of an adult. The recruitment is a
lottery among potential candidates as in the neutral model. The recruitment probability 
is however biased by species specific response to local environmental conditions.

The lottery dynamics described above for the neutral model assume there is a
very large number of offsprings that are candidate for recruitment but only one
will survive and develop to the adult stage. The effect of a differentiation to
local environmental conditions could be implemented at this stage with a biased 
survival probability. The $J_x$ individuals all experience a unique environmental condition 
$E_{nx}$ called a microsite $n$. We considered a patch average $\overline{E_x}$, with a
within-patch variance $\sigma_x$. The regional average is $\overline{E_R}$ and
the regional variance $\sigma_R$ (for simplicity we considered normal
distributions, but different distributions will lead to different regional
similarity constraints \parencite{Mouquet2003,Tilman2004,Gravel2006}). We
consider that a fraction $\lambda_{inx}$ of offsprings reaching the microsite
where recruitment occurs will survive. The recruitment probability is therefore
biased in favour of the species with highest survival because only some species will be able to cope with the local environmental conditions. For tractability we define the relative abundance in
the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. The
calculation of the relative abundance in the seed rain is the same as the
neutral model but the recruitment probability differs because only a fraction of
offspring survive. It is formulated as $Pr_{ix} =
\frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
the relationship between a microsite condition and survival could take various
forms; we used the traditional gaussian curve describing the niche,
$\lambda_{inx} =  \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
niche optimum and $b_i$ is niche breadth. Note that the model will converge to a
neutral model when the niche breadth tends to infinity (which is in fact how we
simulated neutral dynamics in the Supplementary Material to minimize the
complexity of the code).

>>>>

% TODO:
% DG: First complete reading
% DG: Figures
% DG: Legends
% DG: Compile references
% TP & PDP: fix bugs I generated!!!
% DG: Abstract

@@

Ecology needs to move toward a more predictive approach, integrating elements of
theoretical ecology \parencite{Thuiller2013}. The metacommunity perspective \parencite{Leibold2004a}
appears naturally as the appropriate conceptual framework to fill this
challenge. The metacommunity concept builds on feedbacks between local scale
processes, such as competitive interactions and local adaptation, and regional
scale processes such as dispersal, gene flow and speciation. It is particularly relevant to limnology,
where exchanges of organisms and nutrients affect community and ecosystem
properties from the local (e.g. vertical mixing \parencite{Ryabov2011}) to the
regional (e.g. connection of lakes \parencite{Leibold2004b, Gravel2010} scales.
It emphasizes the importance of dispersal relative to pairwise interactions in
the organization of ecological communities.

At first sight, the development of neutral theory appears as a step back.
Neutral theory of biodiversity makes the provocative assumption that species are
ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological communities are
driven only by demographic stochasticity and dispersa and thus, variation in the environment has no impact on
demography. Neutral theory sparked
an historical debate still lasting after more than a decade
\parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It was stimulated
by the surprising ability of neutral models to fit some well studied empirical
observations such as species abundance distributions and distance-decay
relationships.

A remarkable strength of the theory is to provide a \emph{"formal general theory
of abundance and diversity that will account, in a simple and economical
fashion, for the many patterns that ecologists have documented"}
\parencite{Bell2001}. Even if new studies rejecting neutral theory are
consistently published (e.g. \textcite{Ricklefs2012}), a consensus is forming
that neutral theory is a well-developed null hypothesis for niche theory and
could even be used as an adequate approximation of ecological dynamics in some
situations. \textcite{Bell2001} nicely envisioned two perspectives to neutral
theory that are still standing today. Under the weak perspective, neutral theory
provides a set of realistic predictions of community organization despite false
assumptions. Even if being fundamentally wrong, neutral theory would still
useful when used as a null hypothesis \parencite{Gotelli2006}. It is considered
as an improvement over traditional null hypotheses based on randomization
\parencite{Gotelli2000} because it readily integrates dispersal. On the other
had, the strong version posits that neutral theory is a satisfying approximation
to community dynamics and an appropriate theory to explain the distribution of
biodiversity. It implies that the right mechanisms have been identified and that
the consistently observed differences among species do not impact community
organization.

Neutral theory has also been proposed as a useful tool to understand and predict
some aspects of community dynamics. It links to an old philosophical debate
between realism and instrumentalism \parencite{Wennekes2011}. Because every
ecological model is a simplification of reality, scientists have to subjectively
decide the level of details they put in, leaving out some elements they consider
unimportant. The realism perspective requires that all assumptions of the theory
must be true, while the utility of the theory is more important to
instrumentalism. The utilitarian value of a theory could either be for
understanding or for prediction (another old philosophical debate, see
\textcite{Schmueli2010}). Obviously neutral theory could only be instrumental.
The question then is if such a 'general, large-scale, but vague' theory
\parencite{Wennekes2011} is a satisfying approximation.

The instrumentalist view of neutral theory raises the question of why it should
be a satisfying approximation despite knowing the pieces are wrong? Perhaps
stochasticity of various origins blur the deterministic differences among
species and promote ecological drift \parencite{Gravel2011a}. Much has been said
the existence of demographic stochasticity \parencite{Clark2012}, and we
therefore will let this discussion for other papers. A second answer is that
dispersal and historical contingencies might have a much more profound impact on
species distribution \parencite{Bahn2007, Boulangeat2012} and ecological
dynamics. The debate over the equivalence assumption and demographic
stochasticity has perhaps overlook the recognition of how much dispersal
influence community assembly.

@@

A network is a discrete mathematical object made of two sets: a set of nodes
(or vertices) and a set of edges connecting the nodes \cite{new10}. The
term ``graph'' is often preferred in computer science and mathematics
\cite{gro06}, with graph algorithms being an important and active area of
research \cite{sed01}. A network is a combinatorial object: it is used to study
how discrete entities are connected and how they combine together to create
complex structures. They are used to study molecules, food webs, social
networks, or even the relationship between variables in statistics
\cite{wri21,new10}. We are especially interested in spatial networks, a special
kind of network mixing the combinatorial properties of networks with a
topological space \cite{kob94}. Thus, the vertices in a spatial graph are
embedded in some other space, most often the two or three- dimensional Euclidean
space. This object brings a rich representation to spatial ecology and is
particularly suited for systems of lakes and rivers, which can easily be
represented by vertices and edges. There are two notions of distance in spatial
networks. Euclidean distance represents the geographical distance between the
nodes $(i, j)$, i.e.: $\sqrt{(x_i - x_j)^2 + (y_i - y_ j)^2}$. Geodesic
distance is the distance in the graph space, i.e.: the length of the shortest
path \cite{dij59}. For example, two lakes could be very close on a map (short
Euclidean distance) but the geodesic distance could be great if they are not
directly linked by a river.

The popularity of network theory stems for its ability to model complex
structures while allowing us to extract useful metrics (Table 1). At a very high
level, a network can be described by its number of nodes (the order) and 
edges (the size). Looking more closely, the relationship between
nodes is influenced by paths, which are ordered series of nodes. Centrality
is a \emph{central} concept in network theory, where it can be seen
to as a measure of ``importance''. The simplest measure of the centrality of a
node is its degree, which is the number of nodes directly connected to it. Of
course, this is a very rough description of centrality. For example, two lakes
can have the same degree, with one being connected to a small isolated cluster,
while the other one is part of one of the biggest network of lake. In this case,
measures of centrality like eigen-centrality will weight the importance of the
connection, so a node connected to well-connected nodes will have higher
centrality than a node connected to isolated nodes.

For simulations, spatial networks can easily be generated with the random
geometric graph algorithm \cite{sed01}. In this algorithm, all nodes are
assigned to a position in some two-dimensional space, most often the unit
square. Then, all pairs of nodes within some threshold Euclidean distance $r$
are connected with an edge. The resulting networks have the desirable property
of locality: if a node $A$ is connected to two vertices $B$ and $C$. then $B$
and $C$ are more likely to be connected than two random vertices. Random
geometric networks have been extensively studied \cite{app97a,app97
b,app02a,app02b,pen03} and we provide a R function to generate them. The
position of nodes is typically random, but we could also imagine alterations
where they are either more aggregated or segregated than expected by chance
alone.

We also provide the code for a second structure that we call a random geometric
tree. The algorithm first builds a random geometric graph, then select a node
from which to start the tree. It then calculaes the the shortest path tree
\cite{dij59} from this node to all other ones. Edges that are not located along
this tree are pruned. This random geometric tree does not exactly represent
dendritic landscapes but is a convenient model to simulate lake connected by
rivers to a series of smaller lakes.

@@

\begin{table*}
	\centering
	\begin{tabular}{p{5cm}p{8cm}}
		\hline
		\textbf{Concept} & \textbf{Definition} \\
		\hline
		Path & A sequence of edges forming a sequence of nodes \\
		Connection & Two nodes are connected if there is a path between them \\
		Euclidean distance & Geographical distance between two nodes \\
		Geodesic distance & Length of the shortest path between two nodes \\

		\textbf{Network-level metrics} & \\
		Order & Total number of nodes \\
		Size & Total number of edges \\
		Connectivity & A measure of robustness: the minimum number of elements to remove to isolate the nodes \\
		Components &  The number of connected subsets\\

		\textbf{Node-level metrics} & \\
		Degree & The number of edges of a node \\
		Closeness centrality & Average geodesic distance between a node and all other vertex \\
		Eigenvector centrality & A measure of centrality based on the concept that connection to
		highly connected nodes are more important \\
		Betweenness centrality & The number of shortest paths from all nodes to all others that
		pass through that note \\
		\hline
	\end{tabular}
	\caption{Main descriptors of spatial networks used in this study.}
\end{table*}

@@

In this section we describe three toy models representing different
perspectives of metacommunity ecology: patch dynamics, neutral dynamics and
species sorting. While the neutral model is interesting in itself, it is by its
comparison with a model without any interactions (patch dynamics) and with niche
differentiation (species sorting) that we will be able to fully understand the
interaction between these processes and landscape structure. Despite neutral,
competitive interactions in neutral models are very strong because of the
zero-sum assumption (the community is always at carrying capacity). We will
first review the fundamental assumptions of each model with their description
(Table 2 summarizes the parameters and variables that are used), and then
briefly discuss their main predictions. Simulation results are presented in the
next section, with the corresponding R code provided in the Supplementary
Material.

@@

The simplest metacommunity model is a $S$ species extension of traditional
metapopulation models \parencite(Hanski1999). The standard Levins metapopulation
model \parencite(Levins1969) describes the stochastic colonizations and
extinctions of a single species over a homogenous landscape. The basic unit is
the population. The model tracts the dynamics of occupancy (the fraction of the
landscape that is occupied) with an ordinary differential equation and therefore
assumes an infinite landscape. The simulation model we run is more realistic as
it simulates a finite number $N$ of discrete patches (or nodes in network
terminology). The rules described in the previous section were used to generate
connectivity matrices along four scenarios (Fig. 1): global dispersal (connected
graph), a lattice, a random geometric graph and a random tree graph. A patch $x$
shares $d_x$ links with neigbhouring patches (its degree). At each time step
(the simulation model is discrete in time), the probability that a colonist
coming from an occupied patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$
is the probability a colonization event takes place if all connected patches are
occupied. The expected probability that a colonist arrives to patch $x$ from
patch $y$ is then $C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability
that patch $y$ is occupied by species $i$. The probability that an extinction
occurs in a given patch is $e$. The Levins model is for a single species, but a
basic metacommunity patch dynamics model could be run by aggregating $S$
independent metapopulation models \parencite(Hanski1997). There are no
interactions in this simple model, which means there is no limit to local
species richness and no carrying capacity. Competitive, mutualistic and
predator-prey interactions have been added to this framework (e.g.
\textcite{Tilman1994,Holt1996,Klausmeier1998,Gravel2011b}) but we will keep this model
minimal for the sake of comparison with the neutral model.

Predictions of the patch dynamics metacommunity model are quite straightforward.
First, a fundamental result of metapopulation ecology is that persistence will
occur if colonization probability is larger than extinction probability ($c>e$).
Given that all species are the same, then we should expect the regional
diversity ($\gamma$) to be $S$ if this condition is satisfied and $0$ if not. The situation
is however more complex in spatially explicit landscapes with complex
connectivity matrices \parencite{Hanski1998}. Spatially explicit dispersal
usually reduces the occupancy and thereby the likelihood of persistence. The
second prediction is that, given spatial variation in connectivity, there will
be spatial variation in occurrence probability. Given the above formulation of a
colonization event to occur, the probability that an empty location is colonized
is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation basically tells us that the
colonization probability will increase asymptotically with the degree of a patch
(because of the product). It is easy to show from metapopulation theory that the
occurrence probability in a patch is then $p_ix=I_x(I_ix+e)^{-1}$. The feedback
between local and regional dynamics arises because all $p_ix$ from the landscape
are dependent from each other. Simulations are usually conducted to solve the
model for a large landscape, but numerical solutions are theoretically possible.
The aggregation across the $S$ species of the regional species pool is obtained
by taking the summation of occurrence probabilities over all species, $s_x =
\sum{p_i}$. Because in this model all species are equal, we expect the local
species richness to be a linear function of the patch degree (number of edges). Multi-species
analysis of metapopulation models also reveals interesting predictions on other
aspects of community organization at various spatial scales such as the species-area
relationship \parencite{Hanski1997}, and proved to be useful in conservation
ecology with predictions of extinctions following habitat destruction
\parencite{Nee1994,Rybicki2013}.

@@

The formulation of the recruitment probability is the central piece of all
neutral models, making possible the coupling with the metacommunity and
neighbouring patches. We adopt a simple formulation based on
\parencite{Gravel2006}. The approach is conceptually similar to placing a trap
in a canopy gap and picking a seed at random among the ones falling in to
determine the identity of the recruited species. The composition of the seed
pool in that trap will be a mixture of local dispersal and immigrants from the
metacommunity. For simplicity, we consider three spatial scales of dispersal but
it would be easy to generalize the approach to a continuous seed dispersal
kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
recruit is a migrant from neighbouring patches, $M$ is the probability it comes
from a larger (and fixed) metacommunity, and consequently, by substraction,
$1-m-M$ is the probability it comes from local dispersal. The fraction
$N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative
abundance of species $i$ in the seed pool coming from neigbouring patches $x$.
The relative abundance in the neighborhood is weighted by the degree of the
connected nodes because some nodes will spread their seeds across a higher
number of nodes and thus contribute less to the seed pool. We thus consider
$P_{ix} =\frac{\sum P_{iy}d_y^{-1}}{\sum d_y^{-1}}$. We assume for simplicity
(and without loss of generality, \parencite{Bell2000}) that the relative
abundance in the metacommunity is uniform, i.-e. equal to $S^{-1}$. This
immigration prevents the collapse of the metacommunity to a single species,
since otherwise all species except one will face extinction by ecological drift
(speciation prevents this phenomenon to occur in \textcite{Hubbell2001}). The
local recruitment probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
(1-m-M)N_{ix}J_x^{-1}$.

The model is neutral because it assumes that the probabilities of local
recruitment, immigration and mortality events are all equal across species.
Demographic stochasticity is the source of variations in abundance, but larger
disturbances could be simulated as well, as long as they hit all species with
the same probability, independently of their abundance. The fundamental feature of
neutral dynamics is therefore the ecological drift, defined as population
changes emerging from neutraly stable population dynamics. It can be measured as
the variance between replicated time series of community dynamics
\parencite{Gravel2011a}. \textcite{Hubbell2001} provides a very comprehensive
analysis of the model, with specific attention to the effect of the different
parameters on drift (and consequently variance in abundance) and time to
extinction. Despite its simplicity, the neutral model is surprisingly rich in
the predictions it makes. \textcite{Bell2001} and \textcite{Hubbell2001}
analyzed the performance of neutral models to predict species abundance
distributions, the range-abundance relationship, spatial variation in abundance,
the species-area relationship, community turnover (beta-diversity) and
co-occurrence. Recent trophic neutral models were also found to predict realistic ecological
network structures \parencite{Canard2012}. Other than the ecological equivalence
assumption, one of the most criticized aspect of neutral models is the realism
of the speciation process and the required speciation rates to sustain species
richness \parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more
credible speciation models \parencite{Rosindell2009,Desjardins2012a} revealed
the difficulty to maintain diversity in neutral models over macro-evolutionary
time scales. These models nonetheless proposed interesting predictions
on endemic species richness and island biogeography \parencite{Rosindell2011,Desjardins2012b}.

@@

The species-sorting and the mass effect perspectives build on the notion of
species-specific responses to a spatially varying environment
\parencite{Leibold2004a}. There are various ways to simulate such dynamics and
we picked the lottery model, in line with tradition \parencite{Mouquet2002} and
for its proximity to the neutral model described above. Competition for space
occurs during recruitment after the death of an adult. The recruitment is a
lottery among potential candidates as in the neutral model. The recruitment probability 
is however biased by species specific responses to local environmental conditions.

The lottery dynamics described above for the neutral model assume there is a
very large number of offsprings that are candidate for recruitment but only one
will survive and develop to the adult stage. The effect of a differentiation to
local environmental conditions could be implemented at this stage with a biased
survival probability. The $J_x$ individuals all experience a unique
environmental condition $E_{nx}$ called a microsite $n$. We considered a patch
average $\overline{E_x}$, with a within-patch variance $\sigma_x$. The regional
average is $\overline{E_R}$ and the regional variance $\sigma_R$ (for simplicity
we considered normal distributions of environmental conditions, but different
distributions will lead to different regional similarity constraints
\parencite{Mouquet2003,Tilman2004,Gravel2006}). We consider that a fraction
$\lambda_{inx}$ of offsprings reaching the microsite where recruitment occurs
will survive. The recruitment probability is therefore biased in favour of the
species with highest survival because only some species will be able to cope
with the local environmental conditions. We define the relative
abundance in the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} +
(1-m-M)N_{ix}J_x^{-1}$. The calculation of the relative abundance in the seed
rain is the same as the neutral model but the recruitment probability differs
because only a fraction of offspring survive. It is formulated as $Pr_{ix} =
\frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
the relationship between a microsite condition and survival could take various
forms; we used the traditional gaussian curve describing the niche,
$\lambda_{inx} = \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
niche optimum and $b_i$ is niche breadth. Note that the model will converge to a
neutral model when the niche breadth tends to infinity (which is in fact how we
simulated neutral dynamics in the Supplementary Material to minimize the
complexity of the code).

Analyses of similar models with a combination of dispersal and species-sorting
shown that predictions are extremely variables and depend on the frequency
distributions of environmental conditions, niche optimums and breadth. For
instance, a well-studied prediction of neutral models is the species abundance
distribution. It was shown that niche models can predict similar distributions
given appropriate parameters \parencite{Tilman2004,Gravel2006}. The main
prediction is nonetheless that stable and predictable (meaning which species
will coexist) if species are sufficiently dissimilar, which differs from neutral
models. Local species richness will first depend on the joint effects of local
heterogeneity and niche breadth because coexistence requires a sufficient
dissimilaritya mong species \parencite{Schwilk2005}. Local species richness
could be increased by a mass effect when dispersal is consistently supplying
individuals coming from more favorable locations (refuges). The limiting
similarity required to maintain regional coexistence depends on the amount of
dispersal because exchanges among communities homogenizes environmental
conditions. This is one of the main result from the species sorting theory and a
clever example of local-regional feedbacks: increasing dispersal promotes local
coexistence, but on the other hand it diminishes regional coexistence. Only the
best average competitors will remain at very high dispersal. We therefore expect
a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity, with a peak at
intermediate dispersal. On the other hand, we expect a monotonic decrease of
beta ($\beta$) and $\gamma$ diversity with dispersal \parencite{Mouquet2003}. This prediction
has been validated in some experiments \parencite{Venail2008, Logue2011}.
